Millions of Type Errors

### 2023-05-24 oopsla23/paper.tex
rq notes, rw notes

With this telemetry data, we investigate research questions about
the adoption and benefits of type analysis:
\begin{description}
  \item[RQ1.]
    How many creators use type analysis,
    how often does a project use multiple analysis modes,
    and how often do creators turn the analysis off?
  \item[RQ2.]
    For modules that use type analysis: which errors
    arise, how do users respond, and which errors
    tend to persist through subsequent edits?
  \item[RQ3.]
    What impact, if any, does type analysis have on overall code quality?
\end{description}

Answers to these questions can inform the design of gradual languages
in general, and of extensions to \anon{Lua} and languages with semantic
subtyping in particular.

At a higher level, this paper is a step toward data-driven language
design. We are the first to use telemetry as a mechanism for
large-scale measurement of the effectiveness of type error reporting.
Our data captures \textbf{XX} sessions, for a total of \textbf{YY}
hours and \textbf{ZZ} type errors.

Telemetry cannot replace user studies, as there is no way to measure
creator sentiment, but provide complementary data at scale.
FILL can stats answer questions that point-by-point analysis cannot?

@@

@@

>>>>

With this telemetry data, we investigate research questions about
the adoption and benefits of type analysis:
\begin{description}
  \item[RQ1.]
    How many creators use type analysis?
    How often do projects contain modules with different
    analysis modes?
    How often do creators turn analysis off?
  \item[RQ2.]
    For modules that use type analysis:
    which errors arise,
    how do creators respond,
    and which errors tend to persist despite subsequent edits?
  \item[RQ3.]
    What impact does type analysis have on the number of background type
    errors?
    For example, do background errors pile up in unanalyzed projects?
\end{description}

Beyond their immediate applications to \anon{Lua},
answers to these questions have implications for gradual types,
success types, and semantic subtyping.

At a higher level, this paper is the first to use telemetry 
to study a typechecker.
It thus represents a step toward data-driven language design,
informed by many users' actual practice.
Our data captures \textbf{XX} sessions
for a total of \textbf{YY} hours
and \textbf{ZZ} type errors.
By contrast to typical qualitative methods (e.g., surveys and interviews), it
is not restricted to users' \emph{perceptions} about their work and it is not
limited to a small number of users.

FILL can stats answer questions that point-by-point analysis cannot?

@@

Telemetry cannot replace user studies, as there is no way to measure
creator sentiment, but provide complementary data at scale.

@@

\citet{lnsmc-usenix-2018} blame-proportional logging,
scale up ``near'' points of interest,
could be useful for us in the future.
\citet{fnm-sigmod-2020} ditto, faster than statistical at finding root-cause,
deterministic.
